### 2016-09-15 revision/conclusions.tex
typos of reviewer 1

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
regret depends quadratically on norm of the competitor $\|u\|_2$. There
exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
\end{quotation}

>>>>

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
regret depends quadratically on norm of the competitor $\|u\|_2$. There
exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Is there a scale-free algorithm for the decision set $\R^d$, the regret of
which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
\end{quotation}

### 2016-09-15 revision/introduction.tex
typos of reviewer 1

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g. online convex
optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958, Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch and
stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
these domains.

@@

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer.
%Examples of
%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
%projected gradient descent with fixed learning rate.
However, these algorithms have two obvious drawbacks.

@@

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
is vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} have $\Omega(T)$ regret if the Bregman
divergence is not bounded.

@@

For unbounded decision sets no scale-free algorithm with a non-trivial regret
bound was known. Unbounded decision sets are practically important (see e.g.
\cite{Mcmahan-Holt-Sculley-2013}), since learning of large-scale linear models
(e.g. logistic regression) is done by gradient methods that can be reduced to
OLO with decision set $\R^d$.

>>>>

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g., online convex
optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958, Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed bandits
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch and
stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
these domains.

@@

For this reason, most of the existing OLO algorithms receive as an input---or
implicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer.
%Examples of
%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
%projected gradient descent with fixed learning rate.
However, these algorithms have two obvious drawbacks.

@@

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
is vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} might incur $\Omega(T)$ regret if the Bregman
divergence is not bounded.

@@

For unbounded decision sets no scale-free algorithm with a non-trivial regret
bound was known. Unbounded decision sets are practically important (see, e.g.,
\cite{Mcmahan-Holt-Sculley-2013}), since learning of large-scale linear models
(e.g., logistic regression) is done by gradient methods that can be reduced to
OLO with decision set $\R^d$.

### 2016-09-15 revision/preliminaries.tex
typos of reviewer 1

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

>>>>

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable,\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.} the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

### 2016-09-15 revision/solo-ftrl.tex
typos of reviewer 1

In the following, we propose a much simpler and better approach.  We propose to
use Algorithm~\ref{algorithm:ftrl-varying-regularizer} with the regularizer
\begin{equation*}
\label{equation:solo-ftrl-regularizer}
R_t(w) = R(w) \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \; ,
\end{equation*}
where $R:K \to \R$ is any strongly convex function. Through a refined analysis,
we show that the regularizer suffices to obtain an optimal regret bound for any
decision set, bounded or unbounded.  We call such variant \textsc{Scale-free
Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
result is the Theorem~\ref{theorem:regret-solo-ftrl} below, which is proven in
Section~\ref{section:solo-ftrl-regret-bound}.

>>>>

In the following, we propose a much simpler and better approach.  We propose to
use Algorithm~\ref{algorithm:ftrl-varying-regularizer} with the regularizer
\begin{equation}
\label{equation:solo-ftrl-regularizer}
R_t(w) = R(w) \sqrt{\sum_{i=1}^{t-1} \|\ell_i\|_*^2} \; ,
\end{equation}
where $R:K \to \R$ is any strongly convex function. Through a refined analysis,
we show that the regularizer suffices to obtain an optimal regret bound for any
decision set, bounded or unbounded.  We call such variant \textsc{Scale-free
Online Linear Optimization FTRL} algorithm (\textsc{SOLO FTRL}).  Our main
result is Theorem~\ref{theorem:regret-solo-ftrl} below, which is proven in
Section~\ref{section:solo-ftrl-regret-bound}.
